HW 2: Energy Demand and CO2 Emissions

### 2022-09-28 HW2/HW2.Rmd
finish 2:2

tribble(
	~Coefficient, ~Symbol, ~Value,
	"Drag coefficient", "$C_D$", "0.6",
	"Frontal area", "$A_v$", "8.5 m$^2$",
	"Empty vehicle weight", "$m_E$", "11,000 kg",
	"Cargo weight", "$m_C$", "9,000 kg",
	"Rolling resistance coefficient", "$C_R$", "0.006",
	"Density of air", "$\\rho_a$", "1.17 kg/m$^3$",
	"Acceleration of gravity", "$g$", "9.81 m/s$^2$"
) %>% 
	kbl(booktabs = TRUE, escape = FALSE,
			caption = "Assumed Values for Calculations\\label{val}") %>% 
	kable_styling(latex_options = c("HOLD_position"))
```

@@

```{r}
tp <- dc %>% 
	mutate(
		time = `Time (seconds)`,
		v = `Speed (mph)` * 0.44704,
		araw = (v - lag(v)) / (time - lag(time)),
		a = case_when(araw < mina ~ mina, araw > maxa ~ maxa, TRUE ~ araw),
		grade = `Grade (rise/run)`,

@@

		dist = v*(time - lag(time))
		) %>% 
	select(-c(`Time (seconds)`, `Speed (mph)`, `Grade (rise/run)`)) %>% 
	relocate(time, v)
```

@@

## 2.2
```{r}
timemin <- 10000
timemax <- 15000
```

@@

```{r, fig.height=4, fig.cap="Comparison of engine and tractive power.\\label{powercomp}"}
powercomp <- tp %>% 
	filter(time <= timemax, time >= timemin) %>% 
	transmute(time, power, engine = `Engine Power (kW)`,
						diff = engine - power) %>% 
	pivot_longer(-time) %>% 
	mutate(name = as.factor(name))

@@

The data generally seem to make sense, as the engine power is often more than the tractive power, which is expected due to losses. Though I am not sure why tractive power sometimes exceeds engine power; it appears to have something to do with the spikes in tractive power, which are perhaps due to some time gaps.

## 2.3
The total power required to move the vehicle for this drive cycle is `r sum(tp$power, na.rm=TRUE) %>% round()` kW.

## 2.4
```{r}
drag <- sum(tp$dpower, na.rm = TRUE)
roll <- sum(tp$rpower, na.rm = TRUE)
accel <- sum(tp$apower, na.rm = TRUE)
grav <- sum(tp$gpower, na.rm = TRUE)

@@

tibble(
	" " = c("Power (kW)", "%"),
	Drag = c(drag, drag/total * 100),
	Rolling = c(roll, roll/total * 100),
	Acceleration = c(accel, accel/total * 100),
	Gravity = c(grav, grav/total * 100),
	Total = c(total, total/total * 100)
) %>% 
	kbl(booktabs = TRUE, digits = 1, align = 'r',
			caption = "Components of Total Tractive Power\\label{comp}") %>% 
	kable_styling(latex_options = c("HOLD_position"))
```

## 2.5
```{r}
hours <- (max(tp$time) - min(tp$time)) / 3600

regen <- -sum(tp$powerraw[tp$powerraw < 0], na.rm = TRUE)

oldpower <- sum(tp$power, na.rm = TRUE)
newpower <- oldpower - regen
```

Assuming the "negative" tractive power would be used for regenerative braking, an additional `r (regen*hours) %>% round()` kWh would be available (`r regen %>% round()` kW over `r hours %>% round(2)` hours). If 50% of that power was recaptured, then the additional tractive power needed would be reduced to `r newpower %>% round()` kW, a reduction of `r ((oldpower - newpower)/oldpower * 100) %>% round(2)`%.

@@

moveenergy <- sum(enginemove$newengine, na.rm = TRUE)*hours
```

Assuming an auxiliary power load of 7 kW, the engine power used to move the vehicle is $P_{Engine} - 7$ kW, but not less than 0 kW. Finding the total energy expended by the engine attempting to move the vehicle gives `r moveenergy %>% round()` kWh. The total energy actually required to move the vehicle is `r (oldpower*hours) %>% round()` kWh, so the efficiency of the drivetrain is `r (oldpower*hours/moveenergy * 100) %>% round(1)`%.

## 2.7
Low-sulfur diesel has an energy content of 128,488 Btu/gal according to the AFDC. This is equivalent to `r (energy <- 37.656)` kWh/gal. Assuming an engine efficiency of `r (eff <- 0.3)`, the total fuel needed for this drive cycle is $F[\text{gal}] = \frac{`r (oldpower*hours) %>% round()`[\text{kWh}]}{`r energy`[\text{kWh/gal}]}/`r eff` = `r (fuel <- oldpower*hours/energy/eff) %>% round()`\ \text{gal}$. The total distance in this drive cycle is `r (totdist <- sum(tp$dist, na.rm = TRUE)) %>% round()` meters or `r (totdist * .0006213712) %>% round(1)` miles, giving

>>>>

tribble(
	~Coefficient, ~Symbol, ~Value,
	"Drag coefficient", "$C_D$", "0.6",
	"Frontal area", "$A_v$", "8.5 m$^2$",
	"Empty vehicle weight", "$m_E$", "11,000 kg",
	"Cargo weight", "$m_C$", "9,000 kg",
	"Rolling resistance coefficient", "$C_R$", "0.006",
	"Density of air", "$\\rho_a$", "1.17 kg/m$^3$",
	"Acceleration of gravity", "$g$", "9.81 m/s$^2$"
) %>% 
	kbl(booktabs = TRUE, escape = FALSE,
			caption = "Assumed Values for Calculations\\label{val}",
			linesep = "") %>% 
	kable_styling(latex_options = c("HOLD_position"))
```

@@

```{r}
tp <- dc %>% 
	mutate(
		time = `Time (seconds)`,
		t = time - lag(time),
		v = `Speed (mph)` * 0.44704,
		araw = (v - lag(v)) / t,
		a = case_when(araw < mina ~ mina, araw > maxa ~ maxa, TRUE ~ araw),
		grade = `Grade (rise/run)`,

@@

		energy = power*t/3600, #kWh

		dist = v*t
		) %>% 
	select(-c(`Time (seconds)`, `Speed (mph)`, `Grade (rise/run)`)) %>% 
	relocate(time, v)
```

@@

## 2.2
```{r}
timemin <- 13900
timemax <- 14000
```

@@

```{r, fig.height=4, fig.cap="Comparison of engine and tractive power.\\label{powercomp}"}
powercomp <- tp %>% 
	filter(time <= timemax, time >= timemin) %>% 
	transmute(time, power, engine = `Engine Power (kW)`,
						diff = power - engine) %>% 
	pivot_longer(-time) %>% 
	mutate(name = as.factor(name))

@@

The data generally seem to make sense, as the engine power is often more than the tractive power, which is expected due to losses. Though I am not sure why tractive power sometimes exceeds engine power; it appears to have something to do with the spikes in tractive power, which might be due to some time gaps in the data, or possibly a small temporal offset between the engine and tractive power.

## 2.3
The total energy required to move the vehicle for this drive cycle is `r sum(tp$energy, na.rm=TRUE) %>% round()` kWh.

## 2.4
```{r}
drag <- sum(tp$dpower*tp$t/3600, na.rm = TRUE)
roll <- sum(tp$rpower*tp$t/3600, na.rm = TRUE)
accel <- sum(tp$apower*tp$t/3600, na.rm = TRUE)
grav <- sum(tp$gpower*tp$t/3600, na.rm = TRUE)

@@

tibble(
	" " = c("Energy (kWh)", "%"),
	Drag = c(drag, drag/total * 100),
	Rolling = c(roll, roll/total * 100),
	Acceleration = c(accel, accel/total * 100),
	Gravity = c(grav, grav/total * 100),
	Total = c(total, total/total * 100)
) %>% 
	kbl(booktabs = TRUE, digits = 1, align = 'r',
			caption = "Components of Total Tractive Power\\label{comp}") %>% 
	kable_styling(latex_options = c("HOLD_position"))
```

## 2.5
```{r}
regen <- -sum(tp$powerraw[tp$powerraw < 0] * 
								tp$t[tp$powerraw < 0]/3600, na.rm = TRUE)

oldpower <- sum(tp$energy, na.rm = TRUE)
newpower <- oldpower - (regen/2)
```

Assuming the "negative" tractive power would be used for regenerative braking, an additional `r (regen) %>% round()` kWh would be available. If 50% of that power was recaptured, then the additional tractive power needed would be reduced to `r newpower %>% round()` kWh, a reduction of `r ((oldpower - newpower)/oldpower * 100) %>% round(2)`%.

@@

moveenergy <- sum(enginemove$newengine*enginemove$t/3600, na.rm = TRUE)
```

Assuming an auxiliary power load of 7 kW, the engine power used to move the vehicle is $P_{Engine} - 7$ kW, but not less than 0 kW. Finding the total energy expended by the engine attempting to move the vehicle gives `r moveenergy %>% round()` kWh. The total energy actually required to move the vehicle is `r (oldpower) %>% round()` kWh, so the efficiency of the drivetrain is `r (oldpower/moveenergy * 100) %>% round(1)`%.

## 2.7
Low-sulfur diesel has an energy content of 128,488 Btu/gal according to the AFDC. This is equivalent to `r (energy <- 37.656)` kWh/gal. Assuming an engine efficiency of `r (eff <- 0.3)`, the total fuel needed for this drive cycle is $F = \frac{`r (oldpower) %>% round()`[\text{kWh}]/`r eff`}{`r energy %>% round(1)`[\text{kWh/gal}]} = `r (fuel1 <- oldpower/energy/eff) %>% round()`\ \text{gal}$. The total distance in this drive cycle is `r (totdist <- sum(tp$dist, na.rm = TRUE)) %>% round()` meters or `r (miles <- (totdist * .0006213712)) %>% round(1)` miles, giving an overall fuel economy of `r (miles/fuel1) %>% round(2)` mpg, and a fuel consumption of `r (fuel1/miles*1000) %>% round()` gallons per 1000 miles.

## 2.8
The energy required in the regenerative-braking scenario (2.5) is `r newpower %>% round()` kWh. The fuel needed is now $F = \frac{`r (newpower) %>% round()`[\text{kWh}]/`r eff`}{`r energy %>% round(1)`[\text{kWh/gal}]} = `r (fuel2 <- newpower/energy/eff) %>% round()`\ \text{gal}$, with a fuel economy and consumption of `r (miles/fuel2) %>% round(2)` mpg and `r (fuel2/miles*1000) %>% round()` gallons per 1000 miles. If this drive cycle (`r (miles <- (totdist * .0006213712)) %>% round(1)` miles) is representative of a typical day, then at 5 days per week and 52 weeks per year, the fuel saved is $5 \text{[days/wk]} \times 52 \text{[wk/yr]} \times (`r fuel1 %>% round()` - `r fuel2 %>% round()`) \text{[gal]} = `r (fuelsave <- 5*52*(fuel1-fuel2)) %>% round` \text{[gal/yr]}$ for each truck. Given a current (26 Sept. 2022) diesel price of \$`r (price <- 4.889)`/gal, this amounts to a savings of \$`r (price*fuelsave) %>% round()` per year per truck.
